---
title: "P8105 HW2"
author: "Wenbo Fei"
date: "9/29/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

#Problem 1

Read the Mr.Trashwheel dataset
```{r P1_readdata1}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )

head(trashwheel_df)
```

Read the precipitation data from 2018 and 2017
```{r P1_readdata2}
precip_2018 = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation
```{r P1binddata}
precip_df = 
  bind_rows(precip_2018, precip_2017)

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )
  
precip_df = left_join(precip_df, month_df, by = "month")
head(precip_df)
```

The first dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collector collects the trash and store it in a dumpster. This dataset contained detailed information on year, month and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` observations in our final dataset. Each observation has a unique "dumpster" number, while the corresponding year, date, weight of trash(in tons), volume of trash(in cubic yards) and the number some specific kinds of trash will be recorded.

The second dataset includes month precipitation data from year 2018 and 2017. There are a total of `r nrow(precip_df)` observations in our final dataset. Each observation includes the total precipitation corresponding to specific month and year.  

The total precipitation in 2018 is `r sum(pull(precip_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(subset(trashwheel_df, year = 2017), sports_balls))`

